---
title: "Statistical Inference Project - Part One"
author: "Brandon Robinson"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The purpose of the project is to investigate two parts relating statistical analysis of data. Part 1 is to perform a simulation on random data and analyze following the Central Limit Theorem. Part 2 uses a dataset from the R dataset library to perform an analysis and derive a conclusion.

### Part 1 - Simulation Exercise

#### Question 1 - Compare the sample and theoretical mean
First, the required packages will be loaded in order to run the analysis and create the plots

```{r echo = TRUE}
library(dplyr)
library(ggplot2)
```

Next, the base variables will be set in order to run the simulation
```{r echo = TRUE}
#Setting of parameters
set.seed(12345) #provide uniformity of results
lambda <- 0.2
n <- 40
sims <- 1000
```

Then, the simulation will be run in preparation for the plot
```{r echo = TRUE}
simrep <- replicate(sims, rexp(n,lambda))
simmean <- apply(simrep, 2, mean)
theormean <- lambda^-1
```

Finally, the plot of the simulation and theoretical mean will be performed
```{r echo = TRUE}
hist(simmean, col="gray", main = "Sample Mean versus Theoretical Mean", breaks = 20)
abline(v=mean(simmean), lwd = "2", col = "red")
abline(v=mean(theormean), lwd = "2", col = "green")
text(6,100, paste("Actual mean (red) = ", round(mean(simmean),3), "\n Theoretical mean(green) = ",
                   round(theormean,3)))
```

As can be seen in the plot, the value of the real mean and theoretical mean is small and we can assume the Central Limit Theorem is valid in this scenario.

#### Question 2: Compare the sampe and theoretical variances
The next question asks us to compare the variances and note the difference between the real and theoretical
```{r echo = TRUE}
#Sample variance
sampVar <- round(sd(simmean)^2,4)
sampVar
```

```{r echo = TRUE}
#Theoretical variance
theoVar <- round(((1/lambda)/sqrt(n))^2,4)
theoVar
```

The difference between the two variance values is quite small.

#### Question 3: Show the distribution is approximately normal

The final step in the process is to show the distribution is approximately normal.
```{r echo = TRUE}
hist(simmean, prob=TRUE,col="gray", main = "Sample Mean versus Theoretical Mean", breaks = 20)
lines(density(simmean), lwd = 2, col = "green")
x <- rnorm(10000,mean=5,sd = theoVar)
lines(density(x), lwd = 2, col = "red")
```

#### Conclusion
Based on the simulation data and theoretical results, it can be inferred the random distribution is close to normal

